Browsernative Microservices

### 2017-02-27 build/paper.tex
pdf/tex build

\usepackage[normalem]{ulem}
% avoid problems with \sout in headers with hyperref:
\pdfstringdefDisableCommands{\renewcommand{\sout}{}}

@@

Computers can scale, humans can't. Ever since a program or complex
system made by humans has been constrained by humans mental
capabilities. Like in the analogy of juggling balls, our brain can just
``juggle'' a few things at a time. Rich Hickey, the inventor of the
programming language Clojure gave an inspirational keynote on the topic
of \textbf{simplicity}.\footnote{\href{https://www.youtube.com/watch?v=rI8tNMsozo0\&t=46s}{Rails
  Conf 2012 Keynote: Simplicity Matters by Rich Hickey}} In every sphere
of a humans life, simplicity aligns perception with our mental
capacities.

Derived from the Latin word \textbf{simplex}, simple can be understood
as ``literally, uncompounded or onefold''\footnote{\href{http://www.etymonline.com/index.php?term=simple}{Etymology
  Dictionary}} which points directly to the unidimensional aspect. While
complexity describes the multilayered und entangled nature of
conditions, simplicity empowers the human brain to reason about issues
in a straightforward manner. It certainly has some overlapping's with
easy, but while easy is more of a relative nature, simple can be laid
out as a objective manner and therefore universally applicable.

Software development is undoubtedly rich in complexity and full of
subtle pitfalls. In a typical scenario, a piece of software evolves over
time in one or another opinionated direction. Layers of new abstractions
wrestling with old legacy abstractions and mutation becomes untraceable.
Subtle bugs start to creep in. Eventually the small piece of software
may end up in a highly complected monolith which will determine future
design decisions to a painful degree. Future strategies of the
company/organization will be highly determined by the current state in
the need of ``keeping the lights on''.

On the other side of this dystopian scenario, the truly modular system
architecture abandons many of those potential inconsistencies. The whole
system is divided in pluggable parts, object mutation is either
traceable or avoided altogether in favor of immutable data structures.
As Rich Hickey argues, design decisions should be made under the
\textbf{impression of extending, substitution, moving, combining and
repurposing}. The ability to reason about the program at any given time
is crucial for future decisions and implementations. Recalling again the
unidimensional nature of simplicity.

Simplicity in ``the web'', read as a loose generalization of
``everything that runs in the browser'', is certainly a story full of
misconceptions. While simplicity in the backend is mostly a matter of
principles and patterns, any browser-based frontend is restricted on the
highly deterministic nature of the browser platform.

@@

In the recent years many \textbf{frameworks}, libraries and
methodologies approached the global nature of the DOM by scoping assets
and design rules into maintainable components. While the DOM can't be
scoped, JS can. Many frameworks, like ReactJS, AngularJS or VueJS just
to name a few, ditched the old rule of separated HTML, CSS and JS in
favor of an additional layer of abstracted JS components (containing
content, markup and styling). Quiet often those frameworks mimic a MVC
pattern on top of the browser engine which is a reasonable simple design
pattern to build graphical user interfaces. While frameworks are a valid
approach for building scalable web applications they remain highly
opinionated, embody inherent complexity themselves and can change and
break over time. Another downside is code inflation which is a crucial
point for performance. All of those bottlenecks in the web demand for a
new standardized way for creating and evolving complex web services.

@@

JS is the widely accepted language of the web. Nevertheless, a
microservice engineering team might choose another language for various
reasons. Transpiling languages to JS as target language is a stable
solution nowadays. Languages like TypeScript, ClojureScript or
PureScript compile to JS even exclusively. Once web components hit a
critical mass there will be most likely some library support or foreign
function interface towards ES6 modules (which are mandatory for the new
specifications). With the rise of WebAssembly, a new low-level
programming language for the browser, the determination on JS will
hypothetical deteriorate and new quasi native languages for the web
might gain traction.

@@

\subsection{\texorpdfstring{3.1. Custom Elements
\href{http://w3c.github.io/webcomponents/spec/custom/}{(w3c)}}{3.1. Custom Elements (w3c)}}\label{custom-elements-w3c}

@@

\begin{Shaded}
\begin{Highlighting}[]
\OperatorTok{>} \VariableTok{main}\NormalTok{.}\AttributeTok{js}
\KeywordTok{class} \NormalTok{HelloWorld }\KeywordTok{extends} \NormalTok{HTMLElement }\OperatorTok{\{}
 \AttributeTok{constructor}\NormalTok{() }\OperatorTok{\{}
  \KeywordTok{super}\NormalTok{()}\OperatorTok{;} \CommentTok{// mandatory in constructor}
  \KeywordTok{this}\NormalTok{.}\AttributeTok{onclick} \OperatorTok{=} \NormalTok{e }\OperatorTok{=>} \AttributeTok{alert}\NormalTok{(}\StringTok{"hello"}\NormalTok{)}\OperatorTok{;}
 \OperatorTok{\}}
\OperatorTok{\}}
\VariableTok{customElements}\NormalTok{.}\AttributeTok{define}\NormalTok{(}\StringTok{'hello-world'}\OperatorTok{,} \NormalTok{HelloWorld)}
\end{Highlighting}
\end{Shaded}

@@

As mentioned earlier any custom elements must extend the
\texttt{HTMLElement} interface ensuring base properties and methods used
in throughout all HTML elements like \textbf{id, class, addEventListner
\ldots{}}. Additionally, it is possible to define custom attributes
using the \emph{custom elements'} \textbf{getter / setter interface} to
steer the behavior of the element.

\begin{Shaded}
\begin{Highlighting}[]
\OperatorTok{>} \VariableTok{main}\NormalTok{.}\AttributeTok{js}
\KeywordTok{class} \NormalTok{HelloWorld }\KeywordTok{extends} \NormalTok{HTMLElement }\OperatorTok{\{}
  \NormalTok{set }\AttributeTok{sayhello}\NormalTok{(val) }\OperatorTok{\{}
    \KeywordTok{this}\NormalTok{.}\AttributeTok{_hello} \OperatorTok{=} \NormalTok{val}\OperatorTok{;}
    \VariableTok{console}\NormalTok{.}\AttributeTok{log}\NormalTok{(}\KeywordTok{this}\NormalTok{.}\AttributeTok{_hello}\NormalTok{)}\OperatorTok{;}
  \OperatorTok{\}}
  \NormalTok{get }\AttributeTok{sayhello}\NormalTok{() }\OperatorTok{\{}
    \ControlFlowTok{return} \KeywordTok{this}\NormalTok{.}\AttributeTok{_hello}\OperatorTok{;}
  \OperatorTok{\}}
\OperatorTok{\}}
\VariableTok{customElements}\NormalTok{.}\AttributeTok{define}\NormalTok{(}\StringTok{'hello-world'}\OperatorTok{,} \NormalTok{HelloWorld)}\OperatorTok{;}
\CommentTok{// Instantiation via JS instead of HTML}
\KeywordTok{var} \NormalTok{el }\OperatorTok{=} \KeywordTok{new} \AttributeTok{HelloWorld}\NormalTok{()}\OperatorTok{;}
\VariableTok{el}\NormalTok{.}\AttributeTok{sayhello} \OperatorTok{=} \StringTok{"earth"}\OperatorTok{;}
\VariableTok{el}\NormalTok{.}\AttributeTok{sayhello}\OperatorTok{;}\CommentTok{//"earth"}
\end{Highlighting}
\end{Shaded}

@@

\begin{Shaded}
\begin{Highlighting}[]
\OperatorTok{>} \VariableTok{main}\NormalTok{.}\AttributeTok{js}
\KeywordTok{class} \NormalTok{HelloWorld }\KeywordTok{extends} \NormalTok{HTMLElement }\OperatorTok{\{}
 \AttributeTok{constructor}\NormalTok{() }\OperatorTok{\{}
   \KeywordTok{this}\NormalTok{.}\AttributeTok{attachShadow}\NormalTok{(}\OperatorTok{\{}\DataTypeTok{mode}\OperatorTok{:} \StringTok{'open'}\OperatorTok{\}}\NormalTok{)}\OperatorTok{;}
   \VariableTok{shadowRoot}\NormalTok{.}\AttributeTok{innerHTML} \OperatorTok{=} \StringTok{'<p>hello</p>'}\OperatorTok{;}
 \OperatorTok{\}}
\OperatorTok{\}}
\end{Highlighting}
\end{Shaded}

The new global method \texttt{attachShadow} adds a new document root to
the \texttt{HelloWorld} which has the same properties as a normal, light
document object. Note that \texttt{shadowRoot} is marked as
\textbf{open} which ensures that some events can bubble out and outside
JS can reach in the new root.

@@

There is still a missing link between \emph{light DOM} and \emph{shadow
DOM}. The observant reader may have already noticed the weak point in
the HelloWorld example: how to ``vitalize'' the \emph{shadow DOM} with
slot or other element structuring instead of \texttt{innerHTML} strings.
While strings works perfectly fine in this simple case a string of
markup is rather cumbersome and error-prone and doesn't scale well. When
putting quotes inside another quotes things break quickly. It makes the
life hard for developers to work with it because it requires manual
indentation of code blocks and is typically out of syntax highlighting.
That's the time HTML templates come into play.

\subsection{\texorpdfstring{3.3. HTML Templates
\href{https://www.w3.org/TR/html5/scripting-1.html\#the-template-element}{(w3c)}}{3.3. HTML Templates (w3c)}}\label{html-templates-w3c}

Among all other new standards \emph{HTML templates} are the most mature
and adopted standard in the browser environment. All major browsers,
except from Internet Explorer, support this standard.

One core concept in templates is efficiency: Whatever dropped inside a
\texttt{template} tag \sout{bucket} will be parsed on runtime - but not
constructed into the \emph{content tree}. It remains plain HTML Markup
sitting somewhere in the document until the time of activation.

Activation typically takes four steps:

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\tightlist
\item
  \textbf{Querying the template node in question}\\
  const node = document.querySelector(`template');
\item
  \textbf{Parsing the content and preparing the templates' content}\\
  const content = node.content;\\
  -\textgreater{} Returns a \emph{DocumentFragment} object. Handling is
  straight forward content.querySelector(`img').src = `logo.png';
\item
  \textbf{Optional: Cloning the \emph{DocumentFragment} for multiple
  use}\\
  const clone = content.cloneNode(``deep'');
\item
  \textbf{Appending the clone/original to destination}\\
  document.body.appendChild(clone);
\end{enumerate}

As easy and minimal \emph{HTML templates} are, they're missing out a
crucial feature other template implementations usually have. As
templates are basically just dump containers for HTML Markup, there is
no way to define some logic as \textbf{placeholders} where content
should appear. Of course, with heavy use of JS things could be modeled
this way. The idiomatic way tends more towards a \emph{Shadow DOM \&
HTML templates} symbiosis.

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{> index.html}
\KeywordTok{<hello-world>}
 \KeywordTok{<p}\OtherTok{ id=}\StringTok{"sendto"}\OtherTok{ slot=}\StringTok{"placeholder"}\KeywordTok{>}
  \NormalTok{Hello World Web Component  }
 \KeywordTok{</p>}
\KeywordTok{</hello-world>}

\CommentTok{<!-- COMPONENT STARTS HERE -->}
\KeywordTok{<template}\OtherTok{ id=}\StringTok{"hello"}\KeywordTok{>}
 \CommentTok{<!-- STYLES -->}
 \KeywordTok{<style>}
  \FloatTok{#stylewrapper} \KeywordTok{\{}
   \KeywordTok{font-weight:} \DataTypeTok{bold}\KeywordTok{;}
   \KeywordTok{color:} \NormalTok{orange}\KeywordTok{;}
  \KeywordTok{\}}
 \KeywordTok{</style>}
 \CommentTok{<!-- CONTENT -->}
 \KeywordTok{<div}\OtherTok{ id=}\StringTok{"stylewrapper"}\KeywordTok{>}
  \KeywordTok{<slot}\OtherTok{ name=}\StringTok{"placeholder"}\KeywordTok{>}
   \NormalTok{Named placeholder}
  \KeywordTok{</slot>}
 \KeywordTok{</div>}
\KeywordTok{</template>}

\KeywordTok{<script>}
 \CommentTok{// Switched to anonymous class notation}
 \CommentTok{// for keeping associated code together.}
 \VariableTok{customElements}\NormalTok{.}\AttributeTok{define}\NormalTok{(}\StringTok{'hello-world'}\OperatorTok{,}
  \KeywordTok{class} \KeywordTok{extends} \NormalTok{HTMLElement }\OperatorTok{\{}
   \AttributeTok{constructor}\NormalTok{() }\OperatorTok{\{}
    \KeywordTok{super}\NormalTok{()}\OperatorTok{;}
    \KeywordTok{this}\NormalTok{.}\AttributeTok{attachShadow}\NormalTok{(}\OperatorTok{\{}\DataTypeTok{mode}\OperatorTok{:} \StringTok{'open'}\OperatorTok{\}}\NormalTok{)}\OperatorTok{;}
    \KeywordTok{const} \NormalTok{helloTemplate }\OperatorTok{=} \VariableTok{document}\NormalTok{.}\AttributeTok{querySelector}\NormalTok{(}\StringTok{'#hello'}\NormalTok{)}\OperatorTok{;}
    \KeywordTok{this}\NormalTok{.}\VariableTok{shadowRoot}\NormalTok{.}\AttributeTok{appendChild}\NormalTok{(}\VariableTok{helloTemplate}\NormalTok{.}\AttributeTok{content}\NormalTok{)}\OperatorTok{;}
   \OperatorTok{\}}
  \OperatorTok{\}}\NormalTok{)}\OperatorTok{;}
\OperatorTok{<}\SpecialStringTok{/script>}
\end{Highlighting}
\end{Shaded}

The updated \texttt{HelloWorld} component looks already pretty mature.
It combines all the previous mentioned standards into one blob of HTML.
\emph{Custom Elements} serves the logic, \emph{Shadow DOM} scopes the
styles and \emph{HTML Templates} efficiently glues DOM and \emph{Shadow
DOM} together. This separation of concerns comes with a huge gain in
flexibility. In a real world scenario \texttt{HelloWorld} would
contain/reference multiple \emph{HTML Templates} and could switch them
around without any fuss. A developer might to split up templates into
\textbf{STYLE} templates and \textbf{CONTENT} templates to increase
reusability even further.

The last standard in the row of four is not concerned with the internals
of a \emph{web component}. \emph{HTML Imports} serves the need for an
efficient distribution mechanism of components.

\subsection{\texorpdfstring{HTML Imports
\href{https://www.w3.org/TR/html-imports/}{(w3c)}}{HTML Imports (w3c)}}\label{html-imports-w3c}

@@

\begin{Shaded}
\begin{Highlighting}[]
\KeywordTok{<link}\OtherTok{ rel=}\StringTok{"import"}\OtherTok{ href=}\StringTok{"Hello.html"}\OtherTok{ async}\KeywordTok{>}
\end{Highlighting}
\end{Shaded}

The \texttt{async} flag is optional but like in any other fetching
event, strongly recommended. Once the imported HTML document comes into
scope, activation follows a very similar pattern like the aforementioned
\emph{HTML templates}:

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\item
  \textbf{Querying the link node}
\item
  \textbf{Parsing the content and preparing the render}\\
  const content = linknode.import; -\textgreater{} Unlike the \emph{HTML
  template} the content a fully equipped document object.
\item
  \textbf{Optional: Cloning some nodes for multiple use}
\item
  \textbf{Appending the clone/original to destination}
\end{enumerate}

This again is the imperative way to handle a generic \emph{HTML Import}.
In the declarative world of \emph{web components} a component is
activated, parsed and anchored solely by its' tag name
\texttt{\textless{}hello-world\textgreater{}\textless{}/hello-world\textgreater{}}.
Preliminary, the component needs proper configuration. The next section
will elaborate the right configuration and composition of a component to
work out-of-the-box.

Despite from being just a practical document importer \emph{HTML
imports} acts like a fully fledged dependency manager for the browser.
Multiple resources, ranging from stylesheets, scripts, documents, media
files and even other \texttt{imports} can be grouped together in a
logical \texttt{import} statement. Internally, the browser engine keeps
track for every imported resource so it won't be loaded twice. The
inherent complexity is in fact a stumbling block for wider browser
adoption. Currently only Googles blink web engine supports \emph{HTML
Imports} as they are the driving force behind the \emph{web components}
spec in general. Mozilla and Apple imposed distaste for \emph{HTML
Imports} as a whole. One reason for this can be found in the
incompatibility of the spec with the upcoming \emph{ES6 module
loader}.\footnote{https://hacks.mozilla.org/2014/12/mozilla-and-web-components/}

Despite the discrepancies among browser vendors \emph{HTML Imports}
should still be part of the paper as no other native browser technology
can bundle up CSS, JS and HTML that efficient.

\subsection{\texorpdfstring{Appendix A: Custom Events
\href{https://dom.spec.whatwg.org/\#interface-customevent}{(whatwg)}}{Appendix A: Custom Events (whatwg)}}\label{appendix-a-custom-events-whatwg}

Events are first-class citizens in the browser world and \emph{Custom
Events} are no exception. The \emph{Custom Elements} interface is part
of the DOM since years but with the rise of \emph{Custom Elements} they
will most likely become an indispensable building block of \emph{web
components}.

>>>>

@@

Computers can scale, humans can't. Ever since a complex system made by
humans has been constrained by humans mental capabilities. Like in the
analogy of juggling balls our brain can just ``juggle'' a few things at
a time. Rich Hickey, the inventor of the programming language Clojure
gave an inspirational keynote on the topic of
\textbf{simplicity}.\footnote{\href{https://www.youtube.com/watch?v=rI8tNMsozo0\&t=46s}{Rails
  Conf 2012 Keynote: Simplicity Matters by Rich Hickey}} In every sphere
of a humans life simplicity aligns perception with our mental
capacities.

Derived from the Latin word \textbf{simplex}, simple can be understood
as ``literally, uncompounded or onefold''\footnote{\href{http://www.etymonline.com/index.php?term=simple}{Etymology
  Dictionary}} which points towards an unidimensional nature. While
complexity describes the multilayered und entangled nature of
conditions, simplicity empowers the human brain to reason about issues
in a straightforward manner. It certainly has some overlapping's with
easy, but while easy is more of a relative spirit, simple can be laid
out as a objective manner and therefore universally applicable.

Software development is undoubtedly rich in complexity and full of
subtle pitfalls. In a typical scenario a growing software project
evolves in one or another opinionated direction over time. Layers of
abstractions wrestling with aged legacy code requiring layers of
middleware. Mutating assets create subtle bugs and so forth. Eventually
the small piece of software may end up in a highly complected monolith
which will determine future design decisions to a painful degree.
Opaqueness of the system will slow down innovation to a minimum in the
need of ``keeping the lights on''.

On the other side of this dystopian scenario a truly modular system
architecture abandons many of those potential inconsistencies. The whole
system is divided in pluggable parts, object mutation is either
traceable or avoided altogether in favor of immutable data structures.
As Rich Hickey argues, design decisions should be made under the
\textbf{impression of extending, substitution, moving, combining and
repurposing}. The ability to reason about the program at any given time
is crucial for future decisions and implementations. Recalling again the
unidimensional nature of simplicity.

Simplicity in ``the web'', read as a loose generalization of
``everything that runs in the browser'' is certainly a story full of
misconceptions. While simplicity in the backend is mostly a matter of
principles and patterns, any browser-based frontend is restricted on the
highly deterministic nature of the browser platform.

@@

In the recent years many \textbf{frameworks}, libraries and
methodologies approached the global nature of the DOM by scoping assets
and design rules into maintainable components. While the DOM can't be
scoped, JS can. Many frameworks, like ReactJS, AngularJS or VueJS just
to name a few, ditched the old rule of separated HTML, CSS and JS in
favor of an additional layer of abstracted JS components (containing
structure, behavior and sometimes styling). Quiet often those frameworks
mimic a MVC pattern on top of the browser engine which is a reasonable
simple design pattern to build graphical user interfaces. While
frameworks are a valid approach for building scalable web applications
they remain highly opinionated, embody inherent complexity themselves
and can change and break over time. Another downside is code inflation
which is a crucial point for performance. All of those bottlenecks in
the web demand for a new standardized way for creating and evolving
complex web services.

@@

JS is the widely accepted language of the web. Nevertheless, a
microservice engineering team might choose another language for various
reasons. Transpiling languages to JS as target language is a stable
solution nowadays. Languages like TypeScript, ClojureScript or
PureScript compile to JS even exclusively. Once web components hit a
critical mass there will be most likely some library support or foreign
function interface towards ES6 modules (which are mandatory for the new
specifications).

@@

\subsection{\texorpdfstring{3.1. Custom Elements
\href{https://html.spec.whatwg.org/multipage/scripting.html\#custom-elements}{(whatwg)}}{3.1. Custom Elements (whatwg)}}\label{custom-elements-whatwg}

@@

\begin{Shaded}
\begin{Highlighting}[]
\OperatorTok{>} \VariableTok{HelloWorld}\NormalTok{.}\AttributeTok{js}
\KeywordTok{class} \NormalTok{HelloWorld }\KeywordTok{extends} \NormalTok{HTMLElement }\OperatorTok{\{}
 \AttributeTok{constructor}\NormalTok{() }\OperatorTok{\{}
  \KeywordTok{super}\NormalTok{()}\OperatorTok{;} \CommentTok{// mandatory in constructor}
  \KeywordTok{this}\NormalTok{.}\AttributeTok{onclick} \OperatorTok{=} \NormalTok{e }\OperatorTok{=>} \AttributeTok{alert}\NormalTok{(}\StringTok{"hello"}\NormalTok{)}\OperatorTok{;}
 \OperatorTok{\}}
\OperatorTok{\}}
\VariableTok{customElements}\NormalTok{.}\AttributeTok{define}\NormalTok{(}\StringTok{'hello-world'}\OperatorTok{,} \NormalTok{HelloWorld)}
\end{Highlighting}
\end{Shaded}

@@

As mentioned earlier any custom elements must extend the
\texttt{HTMLElement} interface ensuring base properties and methods used
in throughout all HTML elements like \textbf{id, class, addEventListner
\ldots{}}. Additionally, it is possible to define custom attributes
using the \emph{custom elements'} \textbf{getter / setter interface} to
steer the behavior of the element. Note that the keywords \textbf{get}
and \textbf{set} as well as the previously used \textbf{constructor} are
optional!

\begin{Shaded}
\begin{Highlighting}[]
\OperatorTok{>} \VariableTok{HelloWorld}\NormalTok{.}\AttributeTok{js}
\KeywordTok{class} \NormalTok{HelloWorld }\KeywordTok{extends} \NormalTok{HTMLElement }\OperatorTok{\{}
  \NormalTok{set }\AttributeTok{sayhello}\NormalTok{(val) }\OperatorTok{\{}
    \KeywordTok{this}\NormalTok{.}\AttributeTok{_hello} \OperatorTok{=} \NormalTok{val}\OperatorTok{;}
  \OperatorTok{\}}
  \NormalTok{get }\AttributeTok{sayhello}\NormalTok{() }\OperatorTok{\{}
    \ControlFlowTok{return} \KeywordTok{this}\NormalTok{.}\AttributeTok{_hello}\OperatorTok{;}
  \OperatorTok{\}}
\OperatorTok{\}}
\VariableTok{customElements}\NormalTok{.}\AttributeTok{define}\NormalTok{(}\StringTok{'hello-world'}\OperatorTok{,} \NormalTok{HelloWorld)}\OperatorTok{;}
\CommentTok{// Instantiation via JS instead of HTML}
\KeywordTok{var} \NormalTok{el }\OperatorTok{=} \KeywordTok{new} \AttributeTok{HelloWorld}\NormalTok{()}\OperatorTok{;}
\VariableTok{el}\NormalTok{.}\AttributeTok{sayhello} \OperatorTok{=} \StringTok{"earth"}\OperatorTok{;} \CommentTok{// "Call" the setter}
\VariableTok{el}\NormalTok{.}\AttributeTok{sayhello}\OperatorTok{;}\CommentTok{// Yields "earth"}
\end{Highlighting}
\end{Shaded}

@@

\begin{Shaded}
\begin{Highlighting}[]
\OperatorTok{>} \VariableTok{HelloWorld}\NormalTok{.}\AttributeTok{js}
\KeywordTok{class} \NormalTok{HelloWorld }\KeywordTok{extends} \NormalTok{HTMLElement }\OperatorTok{\{}
 \AttributeTok{constructor}\NormalTok{() }\OperatorTok{\{}
   \KeywordTok{this}\NormalTok{.}\AttributeTok{attachShadow}\NormalTok{(}\OperatorTok{\{}\DataTypeTok{mode}\OperatorTok{:} \StringTok{'open'}\OperatorTok{\}}\NormalTok{)}\OperatorTok{;}
   \VariableTok{shadowRoot}\NormalTok{.}\AttributeTok{innerHTML} \OperatorTok{=} \StringTok{'<p>hello</p>'}\OperatorTok{;}
 \OperatorTok{\}}
\OperatorTok{\}}
\end{Highlighting}
\end{Shaded}

The new global method \texttt{attachShadow} adds a new document root to
the \texttt{HelloWorld} which has the same properties as a normal, light
document object. Note that the \texttt{shadowRoot} object is marked as
\textbf{open} which ensures that some events can bubble out and outside
JS can reach in the new root.

@@

There is still a missing link between \emph{light DOM} and \emph{shadow
DOM}. The observant reader may have already noticed the weak point in
the HelloWorld example: how to ``vitalize'' the \emph{shadow DOM} with
slot or other element structuring instead of \texttt{innerHTML} strings.
While strings works perfectly fine in this simple case a string of
markup is rather cumbersome and error-prone and doesn't scale well. When
putting quotes inside other quotes things break quickly. Strings make
development harder because code editor features like indentation or
syntax highlighting won't be supported. The HTML templates is set to
fill this gap.

\subsection{\texorpdfstring{3.3. HTML Templates
\href{https://html.spec.whatwg.org/multipage/scripting.html\#the-template-element}{(whatwg)}}{3.3. HTML Templates (whatwg)}}\label{html-templates-whatwg}

Among all other new specifications \emph{HTML templates} are the most
mature and adopted standard in the browser environment. All major
browsers support it since years.

One core concept in templates is browser performance. Elements inside a
\texttt{template} tag will be parsed on runtime - but not constructed
and rendered into the content tree. They're remaining plain HTML markup
sitting in the document until the time of activation.

Activation usually takes four steps:

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\tightlist
\item
  \textbf{Querying the template node in question}\\
  const node = document.querySelector(`template');
\item
  \textbf{Parsing the content and preparing the templates' content}\\
  const content = node.content;\\
  Returns a DocumentFragment object.
\item
  \textbf{Optional: Cloning the fragment for multiple use}\\
  const clone = content.cloneNode(``deep'');
\item
  \textbf{Appending the clone/original to destination}\\
  document.body.appendChild(clone);
\end{enumerate}

As easy and minimal \emph{HTML templates} are they're missing out a
crucial feature other template implementations used to have. As
templates are basically just containers for HTML markup there is no
idiomatic way to define \textbf{placeholders} for dynamic content.
Templates could be mock up this way using JS for altering the content
but the much cleaner way leverages the previously described slot
technique from shadow DOM.

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{> hello-world-component.html}
\KeywordTok{<hello-world>}
  \KeywordTok{<template}\OtherTok{ id=}\StringTok{"hello"}\KeywordTok{>}
  \CommentTok{<!-- styling -->}
    \KeywordTok{<style>}
      \FloatTok{#hellowrap} \KeywordTok{\{}
        \KeywordTok{font-weight:} \DataTypeTok{bold}\KeywordTok{;}
        \KeywordTok{color:} \NormalTok{orange}\KeywordTok{;}
      \KeywordTok{\}}
    \KeywordTok{</style>}

    \CommentTok{<!-- structure -->}
    \KeywordTok{<div}\OtherTok{ id=}\StringTok{"hellowrap"}\KeywordTok{>}
      \KeywordTok{<slot}\OtherTok{ name=}\StringTok{"placeholder"}\KeywordTok{>}
        \NormalTok{Named placeholder}
      \KeywordTok{</slot>}
    \KeywordTok{</div>}
  \KeywordTok{</template>}

  \CommentTok{<!-- light DOM / typical inserted at index.html -->}
  \KeywordTok{<p}\OtherTok{ slot=}\StringTok{"placeholder"}\KeywordTok{>}
    \NormalTok{Hello World Web Component}
  \KeywordTok{</p>}
\KeywordTok{</hello-world>}

\KeywordTok{<script>}
  \KeywordTok{class} \NormalTok{HelloWorld }\KeywordTok{extends} \NormalTok{HTMLElement }\OperatorTok{\{}
    \AttributeTok{constructor}\NormalTok{() }\OperatorTok{\{}
      \KeywordTok{super}\NormalTok{()}\OperatorTok{;}
      \KeywordTok{this}\NormalTok{.}\AttributeTok{attachShadow}\NormalTok{(}\OperatorTok{\{}\DataTypeTok{mode}\OperatorTok{:} \StringTok{'open'}\OperatorTok{\}}\NormalTok{)}\OperatorTok{;}
      \KeywordTok{const} \NormalTok{temp }\OperatorTok{=} \KeywordTok{this}\NormalTok{.}\AttributeTok{querySelector}\NormalTok{(}\StringTok{'template#hello'}\NormalTok{)}\OperatorTok{;}
      \KeywordTok{this}\NormalTok{.}\VariableTok{shadowRoot}\NormalTok{.}\AttributeTok{appendChild}\NormalTok{(}\VariableTok{temp}\NormalTok{.}\AttributeTok{content}\NormalTok{)}\OperatorTok{;}
    \OperatorTok{\}}
  \OperatorTok{\}}
  \VariableTok{customElements}\NormalTok{.}\AttributeTok{define}\NormalTok{(}\StringTok{'hello-world'}\OperatorTok{,} \NormalTok{HelloWorld)}\OperatorTok{;}
\OperatorTok{<}\SpecialStringTok{/script>}
\end{Highlighting}
\end{Shaded}

The updated \texttt{HelloWorld} component looks already pretty mature.
It combines all the previous mentioned standards into one HTML file.
Custom elements serves the logic, shadow DOM scopes the styles and
\emph{HTML Templates} efficiently glues light DOM and shadow DOM
together. This separation of concerns comes with a surplus in
flexibility. In a real world scenario \texttt{HelloWorld} could
reference multiple \emph{HTML Templates} and switch them around without
any fuss. Even further a developer might split up templates into named
\textbf{STYLE} templates and \textbf{CONTENT} templates to increase
reusability even further.

The last standard in the row of four is not concerned with the internals
implementation of a web component. HTML Imports serves the need for an
efficient distribution mechanism of components and other HTML resources.

\subsection{\texorpdfstring{3.4. HTML Imports
\href{https://www.w3.org/TR/html-imports/}{(w3c)}}{3.4. HTML Imports (w3c)}}\label{html-imports-w3c}

@@

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{> index.html}
\KeywordTok{<link}\OtherTok{ rel=}\StringTok{"import"}\OtherTok{ href=}\StringTok{"Hello.html"}\OtherTok{ async}\KeywordTok{>}

\KeywordTok{<hello-world>}
  \KeywordTok{<p}\OtherTok{ slot=}\StringTok{"placeholder"}\KeywordTok{>}
    \NormalTok{Hello World Web Component}
  \KeywordTok{</p>}
\KeywordTok{</hello-world>}
\end{Highlighting}
\end{Shaded}

The \texttt{async} flag is optional but recommended like any other
fetching event. Once the imported HTML document comes into scope
activation follows a very similar process compared to the aforementioned
HTML templates:

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\item
  \textbf{Querying the link node}
\item
  \textbf{Parsing the content and preparing the render}\\
  const content = linknode.import; -\textgreater{} Unlike the \emph{HTML
  template} a complete document object is constructed.
\item
  \textbf{Optional: Cloning some nodes for multiple use}
\item
  \textbf{Appending the clone/original to destination}
\end{enumerate}

Again this is the imperative way to handle a generic \emph{HTML Import}.
In the declarative world of web components a component is
\textbf{parsed, auto-activated and anchored} solely by its' tag name
\texttt{\textless{}hello-world\textgreater{}\textless{}/hello-world\textgreater{}}
on purpose. The very own lifecycle method \textbf{adoptedCallback()} in
custom elements shows the strong interconnection between those
standards.

\emph{HTML imports} can import everything which is wrappable with HTML
markup. Stylesheets, scripts, documents, media files and even further
imports statements can form a semantic \emph{HTML import} statement. The
far reaching possibilities of a single standard has its downsides.
According to Mozilla \emph{HTML imports} are not fully compatible with
the dependency model of the new ES6 modules.\footnote{https://hacks.mozilla.org/2014/12/mozilla-and-web-components/}
Under the current situation Mozilla and Apple and Microsoft won't
implement \emph{HTML Imports} soon if any. Currently only Googles blink
web engine supports \emph{HTML Imports} as they are the driving force
behind the web components specs in general. Despite the discrepancies
among browser vendors \emph{HTML Imports} are part of this paper as no
other native standard can ship bundled HTML, CSS and JS which is a core
concept behind web components.

\subsection{\texorpdfstring{3.5. Custom Events
\href{https://dom.spec.whatwg.org/\#interface-customevent}{(whatwg)}}{3.5. Custom Events (whatwg)}}\label{custom-events-whatwg}

Events are first-class citizens in the browser. After all they provide a
neat communication channel for dynamic interactions. \emph{Custom
Events} are part of the DOM since years but with the rise of web
components they will most likely become an indispensable building block
of web components.

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{> hello-world-component.html}

\KeywordTok{<hello-world>}
 \KeywordTok{<button>}\NormalTok{Launch CustomEvent}\KeywordTok{</button>}
\KeywordTok{</hello-world>}

\KeywordTok{<script>}
 \VariableTok{customElements}\NormalTok{.}\AttributeTok{define}\NormalTok{(}\StringTok{'hello-world'}\OperatorTok{,} 
 \KeywordTok{class} \KeywordTok{extends} \NormalTok{HTMLElement }\OperatorTok{\{}
  \AttributeTok{constructor}\NormalTok{() }\OperatorTok{\{}
    \KeywordTok{super}\NormalTok{()}\OperatorTok{;}

   \CommentTok{// Craft a helloevent}
    \KeywordTok{const} \NormalTok{helloevent }\OperatorTok{=} \KeywordTok{new} \AttributeTok{CustomEvent}\NormalTok{(}\StringTok{'helloevent'}\OperatorTok{,} \OperatorTok{\{}
    \DataTypeTok{bubbles}\OperatorTok{:} \KeywordTok{true}\OperatorTok{,}
    \DataTypeTok{detail}\OperatorTok{:} \StringTok{'Contains scalar or object'}
   \OperatorTok{\}}\NormalTok{)}\OperatorTok{;}

   \CommentTok{// Launch helloevent if child button clicks}
   \KeywordTok{this}\NormalTok{.}\AttributeTok{addEventListener}\NormalTok{(}\StringTok{'click'}\OperatorTok{,} \NormalTok{click }\OperatorTok{=>} \OperatorTok{\{}
    \KeywordTok{this}\NormalTok{.}\AttributeTok{dispatchEvent}\NormalTok{(helloevent)}
    \VariableTok{click}\NormalTok{.}\AttributeTok{stopPropagation}\NormalTok{()}\OperatorTok{;}
   \OperatorTok{\}}\NormalTok{)}
  \OperatorTok{\}}
 \OperatorTok{\}}\NormalTok{)}\OperatorTok{;}
\OperatorTok{<}\SpecialStringTok{/script>}
\end{Highlighting}
\end{Shaded}

Naming events after the emitting tag makes the API almost
self-explanatory. The \texttt{detail} property can be loaded with
scalars as well as objects. Subsequent parent nodes may catch the custom
event with a clear understanding about the source node.
